### 2017-08-17 Tex_Dissertation/MSc_dissertation.tex
done

The main targets of CLIC are dependent of the energy stage~\cite{clic2016updated}.
In the first it will focus on prescision standard model physics such as Higgs and top quark measurements and in the two subsequent, among others, there will be searches
for new physics. One of the theories that aspires to give solutions to many of the problems of Standard Model is Supersymmetry (SUSY). In this theory every particle has a Supersymmetric partner that differs in the spin by $1/2$, thus 
relating bosons with fermions and fermions with bosons~\cite{bilal2001introduction}.

In the Minimal Supersymmetric Standard Model, the top squarks $\tilde{t}$ can decay into a top quark $t$
and a dark matter candidate, the neutralino $\tilde{\chi}_{1}^{0}$. In this project I will use multivariate analysis
to discriminat between signal and background with the goal to measure the top squark
mass at $\surd{s}$ = 3 TeV in the CLIC accelerator environment.

@@

Specifically, at $\surd s$ = 380 GeV and with an integrated luminosity of $\mathcal{L}_{int}$ = 500 fb$^{-1}$, 
prescision measurements can be made in the Higgs and the top quark sector~\cite{clic2016updated}.
At this enegy stage the Higgsstrahlung
process ($e^{+}e^{-}\rightarrow ZH$) alongside the $WW$ fusion ($e^{+}e^{-}\rightarrow H\nu_{e}\tilde{\nu}_{e}$) 
are the dominant and can shed light to properties of the Higgs boson in a model independent way.
In CLIC there is also a dedicated program for top quark physics. This is because it is one of the most important particles
in SM because it couples strongest to the Higgs field due to its mass but also it has an important role in 
many of the BSM scenarios.
Furthermore at the next two energy stages, investigations will be done with respect to 
proposed scenarios for physics BSM with most 
importantly Supersymmetry at $\surd s$ = 3 TeV and with $\mathcal{L}_{int}$ = 3000 fb$^{-1}$. This  is because CLIC has the potential for direct particle detection up to the 
kinematic limit of $\surd s /2$ for pair production. Also through indirect detection of observables that are sensitive to BSM
scenarios, precision measurements can be made, and comparison with the SM expectations can yield unexpected 
divergencies, 
taking advantage of the full energy potential.

@@

In this chapter I will describe the exact conditions and methods that were followed in this project for the 
measurement of the stop quark mass at the CLIC Conceptual experiment.

@@

In this chapter I describe the procedure that I followed until the calculation of the top squark mass. 
This involved vertex finding, jet clustering, flavour tagging and the calculation of the $\chi^{2}$
for each template of the top squark mass.

@@

As it can be seen the GBDT method achieves higher statistical significance in comparison with the BDT method
but higher uncertainty with a difference of 4 GeV. This can be attributed to the fact that the proposed cuts 
from the TMVA are mainly addressed for particle discoveries rather than for measuring an observable. That 
means that for the optimal precision of the TMVA one has to "play" with the value of the cut -primarily
around the proposed value - in order to extract the lowest uncertainty possible for the mass calculation.
Furthermore as it can be seen in the above table, the number of events after the TMVA cut, in the case of 
GBDT are almost 40 less. Even thought that they sound few, as a fraction they constitute 8$\%$ of the
$N_{signal}$,
number not at all negligible especially when we deal with a few hundreads of events.
Additionally another interesting thing is the importance of these 40 more events in the case of BDT 
(for the mass and the ancertainty calculation).
This is clear because in the BDT the number of events of SM background ($N_{events}$=528) 
in total are far more 
than those in the GBDT method ($N_{events}$=198), and at first sight it would be expected that the GBDT has 
better potential than the BDT.
But the mass and uncertainty measurements proved to be the other way round giving evidence to the fact
that more signal events have more impact than less background.

@@

If the above histogram is compared with the corresponding one for 1000 trees (Figure 4.12), then it is clear
that the SM background has reduced significantly. This is the reason why we get a significance of 11 GeV, i.e.
8 GeV less.

Although, apart from the "tunning" issues, the fact that the BDT's performance is superior is 
something that the experimental community knew already. 
From 2005, similar studies were made for the MiniBoone experiment and 
it was found that BDT's (on Adaboost mode) outperform other methods~\cite{yang2005studies}.

@@

Having in hand the mass of the top squark calculated with the BDT method along with the Monte Carlo mass
of the neutralino, comparison can be made with figure 3.1 of ATLAS. As it can be seen the set of these masses 
($m_{\tilde{t}}$ = 839 GeV/$c^{2}$,$m_{\tilde{\chi}_{1}^{0}}$ = 357 GeV/$c^{2}$) belong in the already 
excluded limits 
of ATLAS experiment. Besides that, the potencial of CLIC collider for a precise measurement of the mass is 
not ruled out but on the opposite stressed, given the potencial of low uncertainty on the mass measurement.

@@

\begin{itemize}
 \item As I investigated the following channel
    \begin{equation}
      e^{-} e^{+} \rightarrow \tilde{t} \tilde{\bar{t}} \rightarrow t\tilde{\chi}_{1}^{0} \bar{t}\tilde{\chi}_{1}^{0}
      \rightarrow W^{+}W^{-}b\bar{b} \tilde{\chi}_{1}^{0} \tilde{\chi}_{1}^{0}
    \end{equation}
the decay of the W boson can be hadronic or leptonic. By isolating and investigating 
the cases that $W\rightarrow l\nu_{l}$
,i.e. the leptonic channel.

@@

\item Besides looking for other decays, work can be done on taking into account on the mass measurement all
the possible stop decays denoted in page 11. In that case, since there will be no SUSY background work has to
be done on simulating a new one involving other interactions.

>>>>

The main targets of CLIC are dependent of the energy stage~\cite{clic2016updated}.
In the first stage it will focus on prescision standard model physics such as Higgs and top quark measurements and in the two subsequent, among others, there will be searches
for new physics. One of the theories that aspires to give solutions to many of the problems of Standard Model is Supersymmetry (SUSY). In this theory every particle has a Supersymmetric partner that differs in the spin by $1/2$, thus 
relating bosons with fermions and fermions with bosons~\cite{bilal2001introduction}.

In the Minimal Supersymmetric Standard Model, the top squarks $\tilde{t}$ can decay into a top quark $t$
and a dark matter candidate, the neutralino $\tilde{\chi}_{1}^{0}$. In this project I will use multivariate analysis
to discriminate between signal and background with the goal to measure the top squark
mass at $\surd{s}$ = 3 TeV in the CLIC accelerator environment.

@@

Specifically, at $\surd s$ = 380 GeV and with an integrated luminosity of $\mathcal{L}_{int}$ = 500 fb$^{-1}$, 
prescision measurements can be made in the Higgs and the top quark sector~\cite{clic2016updated}.
At this enegy stage the Higgsstrahlung
process ($e^{+}e^{-}\rightarrow ZH$) alongside the $WW$ fusion ($e^{+}e^{-}\rightarrow H\nu_{e}\bar{\nu}_{e}$) 
are the dominant and can shed light to properties of the Higgs boson in a model independent way.
In CLIC there is also a dedicated program for top quark physics. This is because it is one of the most important particles
in SM because it couples strongest to the Higgs field due to its mass but also it has an important role in 
many of the BSM scenarios.
Furthermore at the next two energy stages, investigations will be done with respect to 
proposed scenarios for physics BSM with most 
importantly Supersymmetry at $\surd s$ = 3 TeV and with $\mathcal{L}_{int}$ = 3000 fb$^{-1}$. This  is because CLIC has the potential for direct particle detection up to the 
kinematic limit of $\surd s /2$ for pair production. Also through indirect detection of observables that are sensitive to BSM
scenarios, precision measurements can be made, and comparison with the SM expectations can yield unexpected 
divergencies, 
taking advantage of the full energy potential.

@@

In this chapter I will describe the exact conditions and methods that were followed in this project for the 
measurement of the stop quark mass at the CLIC experiment.

@@

In this chapter I describe the procedure that I followed until the measurement of the top squark mass. 
This involved vertex finding, jet clustering, flavour tagging and the calculation of the $\chi^{2}$
for each template of the top squark mass.

@@

As it can be seen the GBDT method achieves higher statistical significance in comparison with the BDT method
but higher uncertainty with a difference of 4 GeV. This can be attributed to the fact that the proposed cuts 
from the TMVA are mainly addressed for particle discoveries rather than for measuring an observable. That 
means that for the optimal precision of the TMVA one has to "play" with the value of the cut -primarily
around the proposed value - in order to extract the lowest uncertainty possible for the mass calculation.
Furthermore as it can be seen in the above table, the number of events after the TMVA cut, in the case of 
GBDT are almost 40 less. Even thought that they sound few, as a fraction they constitute 8$\%$ of the
$N_{signal}$,
number not at all negligible especially when we deal with a few hundreads of events.
Additionally another interesting thing is the importance of these 40 more events in the case of BDT 
(for the mass and the uncertainty calculation).
This is clear because in the BDT the number of events of SM background ($N_{events}$=528) 
in total are far more 
than those in the GBDT method ($N_{events}$=198), and at first sight it would be expected that the GBDT has 
better potential than the BDT.
But the mass and uncertainty measurements proved to be the other way round giving evidence to the fact
that more signal events have more impact than less background.

@@

If the above histogram is compared with the corresponding one for 1000 trees (Figure 4.12), then it is clear
that the SM background has reduced significantly. This is the reason why we get a significance of 11 GeV, i.e.
4 GeV less.

Although, apart from the "tunning" issues, the fact that the BDT's performance is superior is 
something that the experimental community knew. 
From 2005, similar studies were made for the MiniBoone experiment and 
it was found that BDT's (on Adaboost mode) outperform other methods~\cite{yang2005studies}.

@@

Having in hand the mass of the top squark calculated with the BDT method along with the Monte Carlo mass
of the neutralino, comparison can be made with figure 3.1 of ATLAS. As it can be seen the set of these masses 
($m_{\tilde{t}}$ = 839 GeV/$c^{2}$, $m_{\tilde{\chi}_{1}^{0}}$ = 357 GeV/$c^{2}$) belong in the already 
excluded limits 
of ATLAS experiment. Besides that, the potencial of CLIC collider for a precise measurement of the mass is 
not ruled out but on the opposite stressed, given the potencial of low uncertainty on the mass measurement.

@@

\begin{itemize}
 \item As I investigated the following channel
    \begin{equation}
      e^{-} e^{+} \rightarrow \tilde{t} \tilde{\bar{t}} \rightarrow t\tilde{\chi}_{1}^{0} \bar{t}\tilde{\chi}_{1}^{0}
      \rightarrow W^{+}W^{-}b\bar{b} \tilde{\chi}_{1}^{0} \tilde{\chi}_{1}^{0}
    \end{equation}
the decay of the W boson can be hadronic or leptonic. Isolating and investigating 
the cases that $W\rightarrow l\nu_{l}$
,i.e. the leptonic channel a mass measurement can be achieved.

@@

\item Besides looking for other decays, work can be done on taking into account on the mass measurement all
the possible stop decays denoted in page 11. In that case, since there will be no SUSY background
and work has to be done in order to simulate a new one involving other interactions.
